---
title       : SMT201
description : Бие даалтын бодлого
attachments :
  slides_link : https://s3.amazonaws.com/smt201-16/san3.pdf

--- type:NormalExercise lang:r xp:100 skills:1 key:0f9f5045dd
## 15.1 Бодлого

$a_k = frac(\sqrt 2 \cdot k), \quad k=\overline{1,200}$ гишүүнтэй **a** түүврийн хувьд абсалют давтамжаар нь **m=10** хэсэг ангилсан статистикийн цувааг зохио.

*** =instructions
- Өгөгдлийг үзэх,
- `ls()`, `mydata`, `head(mydata)`, `tail(mydata)`, `dim(mydata)`
- Анхны утга олгох,
- `a <- mydata[2]`, `m <- 10`
- ХИУ, ХБУ, алхамыг бодох,
- `amin <- min(a)`, `h <- (amax -amin)/m`
- Баруун үзүүрийн цэг олох,
- `xr < matrix(0,m,1)`
- `for(i in 1:m){ xr[i,1] <- amin + i*h }`
- Төвийн цэгийг олох,
- `xc <- xr - h/2`
- Давтамжийг бодож олох,
- `frq <- matrix(0,m,1)`, `frq[1,1] <- 1`
- `for(i in 1:200){ frq[round(a[i,1]/h),1] <- frq[round(a[i,1]/h),1] +1 }`
- Статистик цувааг зохиох,
- `sseries <- cbind(xc,frq)`

*** =hint

`for(i in 1:200) { frq[round(a[i,1]/h),1] <- frq[round(a[i,1]/h),1] + 1 }`

*** =pre_exercise_code
```{r}
# Хэрэглэгчийн өгөгдөлүүд
# Эдгээр өгөгдөлийг дуудаж ашиглана.
mydata <- read.table("http://s3.amazonaws.com/smt201-16/mydata2.txt")
mydata_csv <- read.csv("http://s3.amazonaws.com/smt201-16/mydata2.csv")
movies <- read.csv("http://s3.amazonaws.com/assets.datacamp.com/course/introduction_to_r/movies.csv")

#library(ggplot2)
#ggplot(movies, aes(x = runtime, y = rating, col = genre)) + geom_point()
```

*** =sample_code
```{r}
# Өгөгдлийг үзэх,
ls();  mydata 
head(mydata)
tail(mydata) 
dim(mydata)

# Анхны утга олгох,
a <- mydata[2]; m <- 10

# ХИУ, ХБУ, алхамыг бодох,
amin <- min(a); h <- (amax -amin)/m

# Баруун үзүүрийн цэг олох,
xr < matrix(0,m,1)
for(i in 1:m){ xr[i,1] <- amin + i*h }

# Төвийн цэгийг олох,
xc <- xr - h/2
#Давтамжийг бодож олох,
frq <- matrix(0,m,1); frq[1,1] <- 1
for(i in 1:m){ frq[round(a[i,1]/h),1] <- frq[round(a[i,1]/h),1] +1 }

#Статистик цувааг зохиох,
sseries <- cbind(xc,frq)
```


*** =sct
```{r}
#no SCT
```

--- type:NormalExercise lang:r xp:100 skills:1 key:ed477d3f70
## More movies

In the previous exercise, you saw a dataset about movies. In this exercise, we'll have a look at yet another dataset about movies!

A dataset with a selection of movies, `movie_selection`, is available in the workspace.

*** =instructions
- Check out the structure of `movie_selection`.
- Select movies with a rating of 5 or higher. Assign the result to `good_movies`.
- Use `plot()` to  plot `good_movies$Run` on the x-axis, `good_movies$Rating` on the y-axis and set `col` to `good_movies$Genre`.

*** =hint
- Use `str()` for the first instruction.
- For the second instruction, you should use `...[movie_selection$Rating >= 5, ]`.
- For the plot, use `plot(x = ..., y = ..., col = ...)`.

*** =pre_exercise_code
```{r}
# You can also prepare your dataset in a specific way in the pre exercise code

library(MindOnStats)
data(Movies)
movie_selection <- Movies[Movies$Genre %in% c("action", "animated", "comedy"),c("Genre", "Rating", "Run")]

# Clean up the environment
rm(Movies)
```

*** =sample_code
```{r}
# movie_selection is available in your workspace

# Check out the structure of movie_selection


# Select movies that have a rating of 5 or higher: good_movies


# Plot Run (i.e. run time) on the x axis, Rating on the y axis, and set the color using Genre

```

*** =solution
```{r}
# movie_selection is available in your workspace

# Check out the structure of movie_selection
str(movie_selection)

# Select movies that have a rating of 5 or higher: good_movies
good_movies <- movie_selection[movie_selection$Rating >= 5, ]

# Plot Run (i.e. run time) on the x axis, Rating on the y axis, and set the color using Genre
plot(good_movies$Run, good_movies$Rating, col = good_movies$Genre)
```

*** =sct
```{r}
# SCT written with testwhat: https://github.com/datacamp/testwhat/wiki

test_function("str", args = "object",
              not_called_msg = "You didn't call `str()`!",
              incorrect_msg = "You didn't call `str(object = ...)` with the correct argument, `object`.")

test_object("good_movies")

test_function("plot", args = "x")
test_function("plot", args = "y")
test_function("plot", args = "col")

test_error()

success_msg("Good work!")
```
